---
title: "Chapter 8 - Data Import with readr2"
date: 2018-02-03
output: github_document
---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(toc = TRUE, toc_depth = 3,comment="#>")
```

```{r prefs, include=FALSE}
options(digits = 3) # personal pref
```


## Introduction

Working with data provided by R packages is a great way to learn the tools of data science, but at some point you want to stop learning and start working with your own data. In this chapter, you'll learn how to read plain-text rectangular files into R. Here, we'll only scratch the surface of data import, but many of the principles will translate to other forms of data. We'll finish with a few pointers to packages that are useful for other types of data.

### Prerequisites

In this chapter, you'll learn how to load flat files in R with the __readr__ package, which is part of the core tidyverse.

```{r setup, message=FALSE}
library(tidyverse)
```

## Getting started

Most of readr's functions are concerned with turning flat files into data frames:

* `read_csv()` reads comma delimited files, `read_csv2()` reads semicolon
  separated files (common in countries where `,` is used as the decimal place),
  `read_tsv()` reads tab delimited files, and `read_delim()` reads in files
  with any delimiter.

* `read_fwf()` reads fixed width files. You can specify fields either by their
  widths with `fwf_widths()` or their position with `fwf_positions()`.
  `read_table()` reads a common variation of fixed width files where columns
  are separated by white space.

* `read_log()` reads Apache style log files. (But also check out
  [webreadr](https://github.com/Ironholds/webreadr) which is built on top
  of `read_log()` and provides many more helpful tools.)

These functions all have similar syntax: once you've mastered one, you can use the others with ease. For the rest of this chapter we'll focus on `read_csv()`. Not only are csv files one of the most common forms of data storage, but once you understand `read_csv()`, you can easily apply your knowledge to all the other functions in readr.

The first argument to `read_csv()` is the most important: it's the path to the file to read.

```{r, eval = FALSE, message = FALSE}
heights <- read_csv("data/heights.csv")
```

When you run `read_csv()` it prints out a column specification that gives the name and type of each column. That's an important part of readr, which we'll come back to in [parsing a file].

You can also supply an inline csv file. This is useful for experimenting with readr and for creating reproducible examples to share with others:

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```


In both cases `read_csv()` uses the first line of the data for the column names, which is a very common convention. There are two cases where you might want to tweak this behaviour:

1.  Sometimes there are a few lines of metadata at the top of the file. You can
    use `skip = n` to skip the first `n` lines; or use `comment = "#"` to drop
    all lines that start with (e.g.) `#`.
    
    ```{r}
    read_csv("The first line of metadata
      The second line of metadata
      x,y,z
      1,2,3", skip = 2)
    
    read_csv("# A comment I want to skip
      x,y,z
      1,2,3", comment = "#")
    ```
    
1.  The data might not have column names. You can use `col_names = FALSE` to
    tell `read_csv()` not to treat the first row as headings, and instead
    label them sequentially from `X1` to `Xn`:
    
    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = FALSE)
    ```
    
    (`"\n"` is a convenient shortcut for adding a new line. You'll learn more
    about it and other types of string escape in [string basics].)
    
    Alternatively you can pass `col_names` a character vector which will be
    used as the column names:
    
    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
    ```
    
Another option that commonly needs tweaking is `na`: this specifies the value (or values) that are used to represent missing values in your file:

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

This is all you need to know to read ~75% of CSV files that you'll encounter in practice. You can also easily adapt what you've learned to read tab separated files with `read_tsv()` and fixed width files with `read_fwf()`. To read in more challenging files, you'll need to learn more about how readr parses each column, turning them into R vectors.

### Compared to base R

If you've used R before, you might wonder why we're not using `read.csv()`. There are a few good reasons to favour readr functions over the base equivalents:

* They are typically much faster (~10x) than their base equivalents.
  Long running jobs have a progress bar, so you can see what's happening. 
  If you're looking for raw speed, try `data.table::fread()`. It doesn't fit 
  quite so well into the tidyverse, but it can be quite a bit faster.

* They produce tibbles, they don't convert character vectors to factors,
  use row names, or munge the column names. These are common sources of
  frustration with the base R functions.

* They are more reproducible. Base R functions inherit some behaviour from
  your operating system and environment variables, so import code that works 
  on your computer might not work on someone else's.

### Exercises

1.  What function would you use to read a file where fields were separated with  
    "|"?
    ```{r eval=FALSE}
    read_delim()
    ```
    
1.  Apart from `file`, `skip`, and `comment`, what other arguments do
    `read_csv()` and `read_tsv()` have in common?
    
    They both have the same arguments, the others being:
    col_types, locale, na, quoted_na, quote, trim_ws, n_max, guess_max, progress
    
    ```{r}
    union(names(formals(read_csv)), names(formals(read_tsv)))
    ```
    
1.  What are the most important arguments to `read_fwf()`?
    Read a fixed width file into a tibble. 
    
    The most important arguments are Column positions, as created by `fwf_empty()`, `fwf_widths()` or `fwf_positions()`.

1.  Sometimes strings in a CSV file contain commas. To prevent them from
    causing problems they need to be surrounded by a quoting character, like
    `"` or `'`. By convention, `read_csv()` assumes that the quoting
    character will be `"`, and if you want to change it you'll need to
    use `read_delim()` instead. What arguments do you need to specify
    to read the following text into a data frame?
    
    ```{r, eval = FALSE}
    "x,y\n1,'a,b'"
    ```
    
    You need to specify `quote = "'"` in:
    
    ```{r}
    read_delim("x,y\n1,'a,b'", delim = ",", quote = "'")
    ```

    or
    
    ```{r}
    read_csv("x,y\n1,'a,b'", quote = "'")
    ```
    
1.  Identify what is wrong with each of the following inline CSV files. 
    What happens when you run the code?
    
    ```{r, eval = FALSE}
    read_csv("a,b\n1,2,3\n4,5,6")
    read_csv("a,b,c\n1,2\n1,2,3,4")
    read_csv("a,b\n\"1")
    read_csv("a,b\n1,2\na,b")
    read_csv("a;b\n1;3")
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n1,2,3\n4,5,6")      # only two columns
    ```

    ```{r warning=TRUE}
    read_csv("a,b,c\n1,2\n1,2,3,4") #(1, 3) is missing and (1, 4) is truncated
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n\"1") # inappropriate extra double quote in string input
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n1,2\na,b") # mismatched types
    ```
    
    ```{r warning=TRUE}
    read_csv("a;b\n1;3") # should use read_csv2() for semi-colon delims
```
    
    