---
title: "Chapter 21 - R Markdown"
date: 2018-02-02
output: html_document
---

# Introduction

R Markdown provides an unified authoring framework for data science, combining your code,
its results, and your prose commentary. R Markdown documents are fully reproducible and
support dozens of output formats, like PDFs, Word files, slideshows, and more.

R Markdown files are designed to be used in three ways:

1. For communicating to decision makers, who want to focus on the conclusions, 
   not the code behind the analysis.
 
2. For collaborating with other data scientists (including future you!), who are interested in 
   both your conclusions, and how you reached them (i.e. the code).
 
3. As an environment in which to do data science, as a modern day lab notebook where you can capture 
   not only what you did, but also what you were thinking.
   
R Markdown integrates a number of R packages and external tools. This means that help is, by-and-large, 
not available through ?. Instead, as you work through this chapter, and use R Markdown in the future, 
keep these resources close to hand:

- R Markdown Cheat Sheet: Help > Cheatsheets > R Markdown Cheat Sheet,
 
- R Markdown Reference Guide: Help > Cheatsheets > R Markdown Reference Guide.
 
Both cheatsheets are also available at http://rstudio.com/cheatsheets.
 
### Prerequisites
 
You need the rmarkdown package, but you don’t need to explicitly install it or load it, 
as RStudio automatically does both when needed.
 
## R Markdown basics
 
This is an R Markdown file, a plain text file that has the extension .Rmd:

<!-- --- -->
<!-- title: "Diamond sizes" -->
<!-- date: 2016-08-25 -->
<!-- output: html_document -->
<!-- --- -->


```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```


It contains three important types of content:

1. An (optional) YAML header surrounded by \-\-\-s.
2. Chunks of R code surrounded by \```.
3. Text mixed with simple text formatting like # heading and \_italics_.

When you open an .Rmd, you get a notebook interface where code and output are interleaved. You can run each code chunk by clicking the Run icon (it looks like a play button at the top of the chunk), or by pressing Cmd/Ctrl + Shift + Enter. RStudio executes the code and displays the results inline with the code:
