---
title: "Chapter 8 - Data Import with readr2"
author: Scott Reeves
# date: 2018-02-03
# output: github_document
output: 
  html_document: 
    toc: true
    toc_float: true
---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(toc = TRUE, toc_depth = 3,comment="#>")
```

```{r prefs, eval=TRUE,include=FALSE}
# options(digits = 3) # personal pref
```


## Introduction

Working with data provided by R packages is a great way to learn the tools of data science, but at some point you want to stop learning and start working with your own data. In this chapter, you'll learn how to read plain-text rectangular files into R. Here, we'll only scratch the surface of data import, but many of the principles will translate to other forms of data. We'll finish with a few pointers to packages that are useful for other types of data.

### Prerequisites

In this chapter, you'll learn how to load flat files in R with the __readr__ package, which is part of the core tidyverse.

```{r setup, message=FALSE}
library(tidyverse)
```

## Getting started

Most of readr's functions are concerned with turning flat files into data frames:

#### read_csv(), read_csv2(), read_tsv(), read_delim()

* `read_csv()` reads comma delimited files, `read_csv2()` reads semicolon
  separated files (common in countries where `,` is used as the decimal place),
  `read_tsv()` reads tab delimited files, and `read_delim()` reads in files
  with any delimiter.
  
#### read_fwf(), fwf_widths(), fwf_positions(), read_table()

* `read_fwf()` reads fixed width files. You can specify fields either by their
  widths with `fwf_widths()` or their position with `fwf_positions()`.
  `read_table()` reads a common variation of fixed width files where columns
  are separated by white space.

#### read_log()

* `read_log()` reads Apache style log files. (But also check out
  [webreadr](https://github.com/Ironholds/webreadr) which is built on top
  of `read_log()` and provides many more helpful tools.)

These functions all have similar syntax: once you've mastered one, you can use the others with ease. For the rest of this chapter we'll focus on `read_csv()`. Not only are csv files one of the most common forms of data storage, but once you understand `read_csv()`, you can easily apply your knowledge to all the other functions in readr.

The first argument to `read_csv()` is the most important: it's the path to the file to read.

```{r, eval = FALSE, message = FALSE}
heights <- read_csv("data/heights.csv")
```

When you run `read_csv()` it prints out a column specification that gives the name and type of each column. That's an important part of readr, which we'll come back to in [parsing a file].

You can also supply an inline csv file. This is useful for experimenting with readr and for creating reproducible examples to share with others:

#### inline csv file

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```


In both cases `read_csv()` uses the first line of the data for the column names, which is a very common convention. There are two cases where you might want to tweak this behaviour:

#### skip = n, comment = "#"

1.  Sometimes there are a few lines of metadata at the top of the file. You can
    use `skip = n` to skip the first `n` lines; or use `comment = "#"` to drop
    all lines that start with (e.g.) `#`.
    
    ```{r}
    read_csv("The first line of metadata
      The second line of metadata
      x,y,z
      1,2,3", skip = 2)
    
    read_csv("# A comment I want to skip
      x,y,z
      1,2,3", comment = "#")
    ```
    
1.  The data might not have column names. You can use `col_names = FALSE` to
    tell `read_csv()` not to treat the first row as headings, and instead
    label them sequentially from `X1` to `Xn`:
    
#### col_names = FALSE
    
    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = FALSE)
    ```
    
(`"\n"` is a convenient shortcut for adding a new line. You'll learn more about it and other types of string escape in [string basics].)
    
Alternatively you can pass `col_names` a character vector which will be used as the column names:
    
    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
    ```
    
Another option that commonly needs tweaking is `na`: this specifies the value (or values) that are used to represent missing values in your file:

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

This is all you need to know to read ~75% of CSV files that you'll encounter in practice. You can also easily adapt what you've learned to read tab separated files with `read_tsv()` and fixed width files with `read_fwf()`. To read in more challenging files, you'll need to learn more about how readr parses each column, turning them into R vectors.

### Compared to base R

If you've used R before, you might wonder why we're not using `read.csv()`. There are a few good reasons to favour readr functions over the base equivalents:

* They are typically much faster (~10x) than their base equivalents.
  Long running jobs have a progress bar, so you can see what's happening. 
  If you're looking for raw speed, try `data.table::fread()`. It doesn't fit 
  quite so well into the tidyverse, but it can be quite a bit faster.

* They produce tibbles, they don't convert character vectors to factors,
  use row names, or munge the column names. These are common sources of
  frustration with the base R functions.

* They are more reproducible. Base R functions inherit some behaviour from
  your operating system and environment variables, so import code that works 
  on your computer might not work on someone else's.

### Exercises

1.  <span style="color:blue">What function would you use to read a file where fields were separated with "|"? </span>
    
    ```{r eval=FALSE}
    read_delim()
    ```
    
1.  <span style="color:blue">Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?</span>
    
    They both have the same arguments, the others being:
    col_types, locale, na, quoted_na, quote, trim_ws, n_max, guess_max, progress
    
    ```{r}
    union(names(formals(read_csv)), names(formals(read_tsv)))
    ```
    
1.  <span style="color:blue">What are the most important arguments to `read_fwf()`?
    Read a fixed width file into a tibble. </span>
    
    The most important arguments are Column positions, as created by `fwf_empty()`, `fwf_widths()` or `fwf_positions()`.

1.  <span style="color:blue">Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like `"` or `'`. By convention, `read_csv()` assumes that the quoting character will be `"`, and if you want to change it you'll need to use  `read_delim()` instead. What arguments do you need to specify to read the following text into a data frame?</span>
    
    ```{r, eval = FALSE}
    "x,y\n1,'a,b'"
    ```
    
    You need to specify `quote = "'"` in:
    
    ```{r}
    read_delim("x,y\n1,'a,b'", delim = ",", quote = "'")
    ```

    or
    
    ```{r}
    read_csv("x,y\n1,'a,b'", quote = "'")
    ```
    
1.  <span style="color:blue">Identify what is wrong with each of the following inline CSV files. 
    What happens when you run the code?</span>
    
    ```{r, eval = FALSE}
    read_csv("a,b\n1,2,3\n4,5,6")
    read_csv("a,b,c\n1,2\n1,2,3,4")
    read_csv("a,b\n\"1")
    read_csv("a,b\n1,2\na,b")
    read_csv("a;b\n1;3")
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n1,2,3\n4,5,6")      # only two columns
    ```

    ```{r warning=TRUE}
    read_csv("a,b,c\n1,2\n1,2,3,4") # (1,3) is missing and (2,4) is truncated
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n\"1") # inline csv file input contains misplaced quote
    ```
    
    ```{r warning=TRUE}
    read_csv("a,b\n1,2\na,b") # mismatched types
    ```
    
    ```{r warning=TRUE}
    read_csv("a;b\n1;3") # should use read_csv2() for semi-colon delims
    ```
    
## Parsing a vector

Before we get into the details of how readr reads files from disk, we need to take a little detour to talk about the `parse_*()` functions. These functions take a __character vector__ and return a more specialised vector like a logical, integer, or date:

#### parse_logical(), parse_integer(), parse_date()

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
```

These functions are useful in their own right, but are also an important building block for readr. Once you've learned how the individual parsers work in this section, we'll circle back and see how they fit together to parse a complete file in the next section.

Like all functions in the tidyverse, the `parse_*()` functions are uniform: the first argument is a character vector to parse, and the `na` argument specifies which strings should be treated as missing:

```{r}
parse_integer(c("1", "231", ".", "456"), na = ".")
```

If parsing fails, you'll get a warning:

```{r warning=TRUE}
x <- parse_integer(c("123", "345", "abc", "123.45"))
```

And the failures will be missing in the output:

```{r warning=TRUE}
x
```

If there are many parsing failures, you'll need to use `problems()` to get the complete set. This returns a tibble, which you can then manipulate with dplyr.

#### problems()

```{r}
problems(x)
```

#### parse_logical(), parse_double(), parse_character(), parse_factor(), parse_datetime(), `parse_date(), parse_time()

Using parsers is mostly a matter of understanding what's available and how they deal with different types of input. There are eight particularly important parsers:

1.  `parse_logical()` and `parse_integer()` parse logicals and integers
    respectively. There's basically nothing that can go wrong with these
    parsers so I won't describe them here further.
    
1.  `parse_double()` is a strict numeric parser, and `parse_number()` 
    is a flexible numeric parser. These are more complicated than you might
    expect because different parts of the world write numbers in different
    ways.
    
1.  `parse_character()` seems so simple that it shouldn't be necessary. But
    one complication makes it quite important: character encodings.

1.  `parse_factor()` create factors, the data structure that R uses to represent
    categorical variables with fixed and known values.

1.  `parse_datetime()`, `parse_date()`, and `parse_time()` allow you to
    parse various date & time specifications. These are the most complicated
    because there are so many different ways of writing dates.

The following sections describe these parsers in more detail.

### Numbers

It seems like it should be straightforward to parse a number, but three problems make it tricky:

1. People write numbers differently in different parts of the world.
   For example, some countries use `.` in between the integer and fractional 
   parts of a real number, while others use `,`.
   
1. Numbers are often surrounded by other characters that provide some
   context, like "$1000" or "10%".

1. Numbers often contain "grouping" characters to make them easier to read, 
   like "1,000,000", and these grouping characters vary around the world.

To address the first problem, readr has the notion of a "locale", an object that specifies parsing options that differ from place to place. When parsing numbers, the most important option is the character you use for the decimal mark. You can override the default value of `.` by creating a new locale and setting the `decimal_mark` argument:

#### parse_double(), parse_double( , locale = (decimal_mark = ","))

```{r}
parse_double("1.23")
parse_double("1,23", locale = locale(decimal_mark = ","))
```

readr's default locale is US-centric, because generally R is US-centric (i.e. the documentation of base R is written in American English). An alternative approach would be to try and guess the defaults from your operating system. This is hard to do well, and, more importantly, makes your code fragile: even if it works on your computer, it might fail when you email it to a colleague in another country.

#### parse_number(), parse_number( , locale = locale(grouping_mark = "."))

`parse_number()` addresses the second problem: it ignores non-numeric characters before and after the number. This is particularly useful for currencies and percentages, but also works to extract numbers embedded in text.

```{r}
parse_number("$100")
parse_number("20%")
parse_number("It cost $123.45")
```

The final problem is addressed by the combination of `parse_number()` and the locale as `parse_number()` will ignore the "grouping mark":

```{r}
# Used in America
parse_number("$123,456,789")

# Used in many parts of Europe
parse_number("123.456.789", locale = locale(grouping_mark = "."))

# Used in Switzerland
parse_number("123'456'789", locale = locale(grouping_mark = "'"))
```

### Strings {#readr-strings}

#### parse_character(), charToRaw()

It seems like `parse_character()` should be really simple --- it could just return its input. Unfortunately life isn't so simple, as there are multiple ways to represent the same string. To understand what's going on, we need to dive into the details of how computers represent strings. In R, we can get at the underlying representation of a string using `charToRaw()`:

```{r}
charToRaw("Hadley")
```

Each hexadecimal number represents a byte of information: `48` is H, `61` is a, and so on. The mapping from hexadecimal number to character is called the encoding, and in this case the encoding is called ASCII. ASCII does a great job of representing English characters, because it's the __American__ Standard Code for Information Interchange.

Things get more complicated for languages other than English. In the early days of computing there were many competing standards for encoding non-English characters, and to correctly interpret a string you needed to know both the values and the encoding. For example, two common encodings are Latin1 (aka ISO-8859-1, used for Western European languages) and Latin2 (aka ISO-8859-2, used for Eastern European languages). In Latin1, the byte `b1` is "±", but in Latin2, it's "ą"! Fortunately, today there is one standard that is supported almost everywhere: UTF-8. UTF-8 can encode just about every character used by humans today, as well as many extra symbols (like emoji!).

readr uses UTF-8 everywhere: it assumes your data is UTF-8 encoded when you read it, and always uses it when writing. This is a good default, but will fail for data produced by older systems that don't understand UTF-8. If this happens to you, your strings will look weird when you print them. Sometimes just one or two characters might be messed up; other times you'll get complete gibberish. For example:

```{r}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"

x1
x2
```

To fix the problem you need to specify the encoding in `parse_character()`:

#### parse_character( , locale = locale(encoding = "Latin1"))

```{r}
parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
```

#### guess_encoding()

How do you find the correct encoding? If you're lucky, it'll be included somewhere in the data documentation. Unfortunately, that's rarely the case, so readr provides  `guess_encoding()` to help you figure it out. It's not foolproof, and it works better when you have lots of text (unlike here), but it's a reasonable place to start. Expect to try a few different encodings before you find the right one.

```{r}
guess_encoding(charToRaw(x1))
guess_encoding(charToRaw(x2))
```

The first argument to `guess_encoding()` can either be a path to a file, or, as in this case, a raw vector (useful if the strings are already in R).

Encodings are a rich and complex topic, and I've only scratched the surface here. If you'd like to learn more I'd recommend reading the detailed explanation at <http://kunststube.net/encoding/>.

### Factors {#readr-factors}

R uses factors to represent categorical variables that have a known set of possible values. Give `parse_factor()` a vector of known `levels` to generate a warning whenever an unexpected value is present:

```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)
```

But if you have many problematic entries, it's often easier to leave as character vectors and then use the tools you'll learn about in [strings] and [factors] to clean them up.

### Dates, date-times, and times {#readr-datetimes}

You pick between three parsers depending on whether you want a date (the number of days since 1970-01-01), a date-time (the number of seconds since midnight 1970-01-01), or a time (the number of seconds since midnight). When called without any additional arguments:

*   `parse_datetime()` expects an ISO8601 date-time. ISO8601 is an
    international standard in which the components of a date are
    organised from biggest to smallest: year, month, day, hour, minute, 
    second.
    
    ```{r}
    parse_datetime("2010-10-01T2010")
    # If time is omitted, it will be set to midnight
    parse_datetime("20101010")
    ```
    
    This is the most important date/time standard, and if you work with
    dates and times frequently, I recommend reading
    <https://en.wikipedia.org/wiki/ISO_8601>
    
*   `parse_date()` expects a four digit year, a `-` or `/`, the month, a `-` 
    or `/`, then the day:
    
    ```{r}
    parse_date("2010-10-01")
    ```

*   `parse_time()` expects the hour, `:`, minutes, optionally `:` and seconds, 
    and an optional am/pm specifier:
    
#### library(hms)
  
    ```{r}
    library(hms)
    parse_time("01:10 am")
    parse_time("20:10:01")
    ```
    
    Base R doesn't have a great built in class for time data, so we use 
    the one provided in the hms package.
    
If these defaults don't work for your data you can supply your own date-time `format`, built up of the following pieces:

Year
: `%Y` (4 digits). 
: `%y` (2 digits); 00-69 -> 2000-2069, 70-99 -> 1970-1999.

Month
: `%m` (2 digits).
: `%b` (abbreviated name, like "Jan").
: `%B` (full name, "January").

Day
: `%d` (2 digits).
: `%e` (optional leading space).

Time
: `%H` 0-23 hour.
: `%I` 0-12, must be used with `%p`.
: `%p` AM/PM indicator.
: `%M` minutes.
: `%S` integer seconds.
: `%OS` real seconds. 
: `%Z` Time zone (as name, e.g. `America/Chicago`). Beware of abbreviations:
  if you're American, note that "EST" is a Canadian time zone that does not
  have daylight savings time. It is _not_ Eastern Standard Time! We'll
  come back to this [time zones].
: `%z` (as offset from UTC, e.g. `+0800`). 

Non-digits
: `%.` skips one non-digit character.
: `%*` skips any number of non-digits.

The best way to figure out the correct format is to create a few examples in a character vector, and test with one of the parsing functions. For example:

```{r}
parse_date("01/02/15", "%m/%d/%y")
parse_date("01/02/15", "%d/%m/%y")
parse_date("01/02/15", "%y/%m/%d")
```

If you're using `%b` or `%B` with non-English month names, you'll need to set the  `lang` argument to `locale()`. See the list of built-in languages in `date_names_langs()`, or if your language is not already included, create your own with `date_names()`.

```{r}
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

### Exercises

1.  <span style="color:blue">What are the most important arguments to `locale()`? </span>

    When parsing numbers, the most important arguments to `locale()` are `decimal mark` and grouping_mark. 
    
1.  <span style="color:blue">What happens if you try and set `decimal_mark` and `grouping_mark` 
    to the same character? What happens to the default value of 
    `grouping_mark` when you set `decimal_mark` to ","? What happens
    to the default value of `decimal_mark` when you set the `grouping_mark`
    to "."?</span>

    ```{r}
    # When `decimal_mark` and `grouping_mark` to the same character the fractional
    # digits become part of the whole
    # 
    parse_number("123,456,789,28", locale = locale(grouping_mark = ","), (decimal_mark = ","))
    ```
    
    ```{r}
    # If `decimal_mark` is set to ",", then `grouping_mark` is superceded.
    # 
    parse_number("123,456,789,28", locale = locale(decimal_mark = ","))
    ```

    ```{r}
    # If `grouping_mark` is set to ".", then `decimal_mark` is superceded.
    # 
    parse_number("123,456,789.28", locale = locale(grouping_mark = "."))
    ```

1.  I didn't discuss the `date_format` and `time_format` options to
    `locale()`. What do they do? Construct an example that shows when 
    they might be useful.

1.  If you live outside the US, create a new locale object that encapsulates 
    the settings for the types of file you read most commonly.
    
1.  What's the difference between `read_csv()` and `read_csv2()`?
    
1.  What are the most common encodings used in Europe? What are the
    most common encodings used in Asia? Do some googling to find out.

1.  Generate the correct format string to parse each of the following 
    dates and times:
    
    ```{r}
    d1 <- "January 1, 2010"
    d2 <- "2015-Mar-07"
    d3 <- "06-Jun-2017"
    d4 <- c("August 19 (2015)", "July 1 (2015)")
    d5 <- "12/30/14" # Dec 30, 2014
    t1 <- "1705"
    t2 <- "11:15:10.12 PM"
    ```
