---
title: "Exercises - Chapter 1 - Data Visualization with ggplot2"
#date: 2018-01-16
author: Scott Reeves
output: 
  html_document: 
    toc: true
    toc_float: true
---

<!-- ```{r setup, include=FALSE} -->
<!-- knitr::opts_chunk$set(echo = TRUE) -->
<!-- ``` -->

```{r setup, include=FALSE}
set.seed(1014)
#options(digits = 3)

knitr::opts_chunk$set(comment = "#>",collapse = TRUE,cache = TRUE,out.width = "70%",fig.align = 'center',fig.width = 6,fig.asp = 0.618)

options(dplyr.print_min = 6, dplyr.print_max = 6)
```



```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
```

## First steps
### Exercises

1.  <span style="color:blue">Run `ggplot(data = mpg)`. What do you see?</span>

    ```{r eval=FALSE}
    ggplot(data = mpg) #An empty graph / a gray box / an empty coordinate system
    ```

2.  <span style="color:blue">How many rows are in `mpg`? How many columns?</span>

    ```{r}
    dim(mpg)
    ```

3.  <span style="color:blue">What does the `drv` variable describe?  Read the help for `?mpg` to find
    out.</span>
    
    ```{r}
    ?mpg
    
    # drv
    # f = front-wheel drive, r = rear wheel drive, 4 = 4wd</span>
    ```

     
4.  <span style="color:blue">Make a scatterplot of `hwy` vs `cyl`.
    ```{r}
    ggplot(mpg) +
      geom_point(aes(cyl, hwy))
    ```


5.  <span style="color:blue">What happens if you make a scatterplot of `class` vs `drv`? Why is
    the plot not useful?</span>
    
    ```{r}
    ggplot(mpg) +
      geom_point(aes(drv, class)) # Informs of existence, nothing more.
    ```


## Aesthetic mappings
### Exercises

1.  <span style="color:blue">What's gone wrong with this code? Why are the points not blue?</span>

    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
    ```
    
    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
    ```


1.  <span style="color:blue">Which variables in `mpg` are categorical? Which variables are continuous? 
    (Hint: type `?mpg` to read the documentation for the dataset). How
    can you see this information when you run `mpg`?</span>
    
    ```{r }
    str(mpg) # all vars are categorical except cty and hwy
    ```
    

1.  <span style="color:blue">Map a continuous variable to `color`, `size`, and `shape`. How do
    these aesthetics behave differently for categorical vs. continuous
    variables?</span>
    
    ```{r}
    ggplot(mpg) +
      geom_point(aes(displ, cty, color = hwy))
    ```
    ```{r}
    ggplot(mpg) +
      geom_point(aes(displ, cty, size = hwy))
    ```
    
    ```{r eval=FALSE}
    ggplot(mpg) +
      geom_point(aes(displ, cty, shape = hwy)) #Will not work. A continuous var cannot be mapped to a shape.
    ```

1.  <span style="color:blue">What happens if you map the same variable to multiple aesthetics?</span> 

1.  <span style="color:blue">What does the `stroke` aesthetic do? What shapes does it work with?</span>
    (Hint: use `?geom_point`)
    
1.  <span style="color:blue">What happens if you map an aesthetic to something other than a variable 
    name, like `aes(colour = displ < 5)`?</span>


## Facets
### Exercises

1.  <span style="color:blue">What happens if you facet on a continuous variable?</span>

1.  <span style="color:blue">What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean?
    How do they relate to this plot?</span>
    
    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = drv, y = cyl))
    ```

1.  <span style="color:blue">What plots does the following code make? What does `.` do?</span>

    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)
    
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
    ```

1.  <span style="color:blue">Take the first faceted plot in this section:</span>

    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
    ```
    
    <span style="color:blue">What are the advantages to using faceting instead of the colour aesthetic?
    What are the disadvantages? How might the balance change if you had a larger dataset?</span>
    
1.  <span style="color:blue">Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other
    options control the layout of the individual panels? Why doesn't
    `facet_grid()` have `nrow` and `ncol` argument?</span>

1.  <span style="color:blue">When using `facet_grid()` you should usually put the variable with more
    unique levels in the columns. Why?</span>

## Geometric objects
### Exercises

1.  <span style="color:blue">What geom would you use to draw a line chart? A boxplot? 
    A histogram? An area chart?</span>

1.  <span style="color:blue">Run this code in your head and predict what the output will look like.
    Then, run the code in R and check your predictions.</span>
    
    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```

1.  <span style="color:blue">What does `show.legend = FALSE` do?  What happens if you remove it?  
    Why do you think I used it earlier in the chapter?</span>

1.  <span style="color:blue">What does the `se` argument to `geom_smooth()` do?</span>


1.  <span style="color:blue">Will these two graphs look different? Why/why not?</span>

    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth()
    
    ggplot() + 
      geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
    ```

1.  <span style="color:blue">Recreate the R code necessary to generate the following graphs.</span>
    
    ```{r echo=FALSE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```
    
    ```{r echo=TRUE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(aes(group = drv), se = FALSE) +
      geom_point()
    ```
    
    ```{r echo=TRUE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```
    
    ```{r echo=TRUE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) + 
      geom_smooth(se = FALSE)
    ```
    
    ```{r echo=TRUE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) +
      geom_smooth(aes(linetype = drv), se = FALSE)
    ```
    
    ```{r echo=TRUE, fig.align="default", fig.width=3, message=FALSE, out.width="50%"}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(size = 4, colour = "white") + 
      geom_point(aes(colour = drv))
    ```

## Statistical transformations

### Exercises

1.  <span style="color:blue">What is the default geom associated with `stat_summary()`? How could
    you rewrite the previous plot to use that geom function instead of the 
    stat function?</span>

1.  <span style="color:blue">What does `geom_col()` do? How is it different to `geom_bar()`?</span>

1.  <span style="color:blue">Most geoms and stats come in pairs that are almost always used in 
    concert. Read through the documentation and make a list of all the 
    pairs. What do they have in common?</span>

1.  <span style="color:blue">What variables does `stat_smooth()` compute? What parameters control
    its behaviour?</span>

1.  <span style="color:blue">In our proportion bar chart, we need to set `group = 1`. Why? In other
    words what is the problem with these two graphs?</span>
    
    ```{r, eval = FALSE}
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, y = ..prop..))
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
    ```
  

## Position adjustments

### Exercises

1.  <span style="color:blue">What is the problem with this plot? How could you improve it?</span>

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
      geom_point()
    ```

1.  <span style="color:blue">What parameters to `geom_jitter()` control the amount of jittering?</span>

1.  <span style="color:blue">Compare and contrast `geom_jitter()` with `geom_count()`.</span>

1.  <span style="color:blue">What's the default position adjustment for `geom_boxplot()`? Create
    a visualisation of the `mpg` dataset that demonstrates it.</span>

## Coordinate systems

### Exercises

1.  <span style="color:blue">Turn a stacked bar chart into a pie chart using `coord_polar()`.</span>

1.  <span style="color:blue">What does `labs()` do? Read the documentation.</span>

1.  <span style="color:blue">What's the difference between `coord_quickmap()` and `coord_map()`?</span>

1.  <span style="color:blue">What does the plot below tell you about the relationship between city
    and highway mpg? Why is `coord_fixed()` important? What does `geom_abline()` do?</span>
    
    ```{r, fig.asp = 1, out.width = "50%"}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() + 
      geom_abline() +
      coord_fixed()
    ```




