---
title: "Chapter 1 - Data Visualization with ggplot2"
date: 2018-01-16
output: html_document
---

<!-- ```{r setup, include=FALSE} -->
<!-- knitr::opts_chunk$set(echo = TRUE) -->
<!-- ``` -->

# Introduction

>“The simple graph has brought more information to the data analyst’s mind than any other device.” — John Tukey

This chapter will teach you how to visualise your data using ggplot2. R has several systems for making graphs, but ggplot2 is one of the most elegant and most versatile. ggplot2 implements the **grammar of graphics**, a coherent system for describing and building graphs. With ggplot2, you can do more faster by learning one system and applying it in many places.

If you’d like to learn more about the theoretical underpinnings of ggplot2 before you start, I’d recommend reading “The Layered Grammar of Graphics”, http://vita.had.co.nz/papers/layered-grammar.pdf.

## Prerequisites

This chapter focusses on ggplot2, one of the core members of the tidyverse. To access the datasets, help pages, and functions that we will use in this chapter, load the tidyverse by running this code:

```{r, echo=TRUE}
library(tidyverse)
```

That one line of code loads the core tidyverse; packages which you will use in almost every data analysis. It also tells you which functions from the tidyverse conflict with functions in base R (or from other packages you might have loaded).

If you run this code and get the error message “there is no package called ‘tidyverse’”, you’ll need to first install it, then run `library()` once again.


# First Steps

Let’s use our first graph to answer a question: Do cars with big engines use more fuel than cars with small engines? You probably already have an answer, but try to make your answer precise. What does the relationship between engine size and fuel efficiency look like? Is it positive? Negative? Linear? Nonlinear?

## The `mpg` data frame

You can test your answer with the `mpg` **data frame** found in ggplot2 (aka `ggplot2::mpg`). A data frame is a rectangular collection of variables (in the columns) and observations (in the rows). mpg contains observations collected by the US Environment Protection Agency on 38 models of car.

```{r}
mpg
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

